---
title: "chapter6"
output: html_document
date: "2022-12-11"
---

```{r}

BPRSL
RATSL

##Implement the analyses of Chapter 8 of MABS, using the R codes of Exercise Set 6: Meet and Repeat: PART I but using the RATS data (from Chapter 9 and Meet and Repeat: PART II).
#(0-7 points: 0-4 points for graphs or analysis results + 0-3 points for their interpretations)

ggplot(RATSL, aes(x = Time, y = Weight, color = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))



```
Observing the resulting graph, we can deduce that groups 2 and 3 have had a significant increase in weight, ranging up to twice as heavy as their counterparts in Group 1.

```{r}


RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stdWeight = Weight) %>%
  ungroup()

glimpse(RATSL)


library(ggplot2)
ggplot(RATSL, aes(x = Time, y = stdWeight, color = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized Weight")
```





```{r}

RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise(mean = mean(Weight), se = sd(Weight)) %>%
  ungroup()


glimpse(RATSS)


ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")


```
Looking at the graph of the mean Weight values, we can still see a clear difference between Groups 2 and 3 respective of the rats in Group 1. We also observe how the difference in average weight for groups 2 and 3 starts by being bigger but over time it becomes much closer. All groups see a steady increase in average weight over time.


```{r}
# Create a summary data by treatment and subject with mean as the summary variable (ignoring baseline week 0)
RATSL8S <- RATSL %>%
  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSL8S)

# Draw a boxplot of the mean versus treatment
library(ggplot2)
ggplot(RATSL8S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), Time 1-8")

# Although it is not needed, we can produce a similar graph to exercise 6 by filtering the outlier with significantly more weight in Group 2
RATSL8S1 <- RATSL8S %>%
  filter(mean < 590) 

ggplot(RATSL8S1, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(bprs), weeks 1-8")
```

By removing the outliar from Group 2, we see how the se of the average weight values in the group is reduced. However, this case is nearly not as clear as it was in the BPRS exercise, so there is not as much of a need to observe the graphs without it.
Simmilarly to the previous parts of analysis, this points to a clear effect of the Group the rats were placed in with the gain in weight over the same time span.

Since the two-sample t-test should be carried with a grouping factor of 2, and the RATS data frame having 3 different groups, it is not possible to replicate it as in Exercise 6.


PART II:

```{r}
dim(BPRSL)

ggplot(BPRSL, aes(x = week, y = bprs, group = treatment)) +
  geom_smooth()
```
By adding "geom_smooth", we avoid finding a zigzagging graph derived of how ggplot connects points that are two close together, and it allows us to see a much more insightful graph.


```{r}

#Next, creating a regression model


BPRS_reg <- lm(bprs ~ week + treatment, data = BPRSL)

# print out a summary of the model
summary(BPRS_reg)
```


```{r}
#Next, the random intercept and random slope exercise

library(lme4)

BPRS_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)

summary(BPRS_ref)



BPRS_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)

summary(BPRS_ref1)

# perform an ANOVA test on the two models
anova(BPRS_ref1, BPRS_ref)

```
The analysis provides a p value of 0.02, which is significantly higher than the Chi-Square value, which points to a better fit of the second model compared to the first one.



```{r}

BPRS_ref2 <- lmer(bprs ~ week + treatment + week:treatment + (week | subject), data = BPRSL, REML = FALSE)

# print a summary of the model

summary(BPRS_ref2)
# perform an ANOVA test on the two models
anova(BPRS_ref2, BPRS_ref1)

# draw the plot of RATSL with the observed Weight values
ggplot(BPRSL, aes(x = week, y = bprs, group = treatment)) +
  geom_smooth(aes(color = subject)) +
  scale_x_continuous(name = "Week") +
  scale_y_continuous(name = "Observed BPRS") +
  theme(legend.position = "top")

# Create a vector of the fitted values
fitted_values <- fitted(BPRS_ref2)

library(dplyr)
library(tidyr)
# Create a new column fitted to RATSL
BPRSL_w_fitted <- BPRSL %>%
  mutate(Fitted = fitted_values)

# draw the plot of RATSL with the Fitted values of weight
library(ggplot2)
ggplot(BPRSL, aes(x = week, y = fitted_values, group = treatment)) +
  geom_smooth(aes(color = treatment)) +
  scale_x_continuous(name = "WEEK") +
  scale_y_continuous(name = "Fitted BPRS") +
  theme(legend.position = "top")



```

In this last analysis, I also used "geom_smooth" to get over the sharp jumps in the graph. It shows how both groups experience a similar slope, but it is important to note how for subjects in group 2 the treatment was less effective, as the average score starded lower than the other group's starting point, and by the last week their avergae is higher than their counterparts with different treatment.


